---
name: Joshua
surname: Rosenberg
position: "Assistant Professor of STEM Education"
address: "1328 Beacon Hill Ln., Knoxville, TN 37919"
phone: "+1 865-236-1445"
www: joshuarosenberg.com
email: "jmrosen48@gmail.com"
twitter: jrosenberg6432
github: jrosen48
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::moderncv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tidyverse)
```

# Highlights

- Established research agenda at the intersection of computational and machine learning methods and social science (particularly educational research)
- Experienced data scientist and developer of statistical software for data analysts, educators, and learners
- Principal Investigator for more than 4M in federally-funded research projects
- Author of a book on educational data science and 37 peer-reviewed journal articles
- Mentor to and supervisor for students in computer science, data science, and education

# Professional Affiliation

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Assistant Professor of STEM Education", "2018-Present", "University of Tennessee", "Knoxville, TN",
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Education

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "PhD, Educational Psychology & Educational Technology", "2018", "Michigan State University", "East Lansing, MI",
  "MA, Educational Technology", "2012", "Michigan State University", "East Lansing, MI",
  "BS, Biology", "2010", "University of North Carolina, Asheville", "Asheville, NC",
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Awards and Fellowships (Selected)

```{r}
tribble(
  ~ Award, ~ Year, ~ Institution, ~ Where,
  "Early Career Award", "2022", "Technology as an Agent of Change for Teaching and Learning Special Interest Group, AERA",
  "Best Poster Award", "2021", "Best Poster Award", "International Conference on Educational Data Mining",
  "Fellowship", "2021-2022", "Open Educational Resources (OER) Research", "William and Flora Hewlett Foundation"
) %>% 
  detailed_entries(Award, Year, Institution, Where)
```

# Grants (Selected)

```{r}
tribble(
  ~Role,  ~Years, ~ Title, ~Funder,
    "Co-PI", "2022-2025", "Quantifying the robustness of causal inferences: Extensions and applications", "Institute for Educaiton Sciences ($899,319)",
  "PI","2019-2022",  "Understanding the development of interest in computer science: An experience sampling approach", "NSF ($348,688)",
  "Co-PI", "2019-2022", "Advancing computational grounded theory for audiovisual data from STEM classrooms", "NSF ($1,313,855)"
) %>% 
    detailed_entries(Role, Years,  Title, Funder)
```

# Publications (Selected)

```{r}
bibliography_entries("publications.bib") %>%
  arrange(issued)
```

# Software Developed (Selected)

```{r}
bibliography_entries("software.bib") %>%
  arrange(issued)
```

