---
name: Joshua
surname: Rosenberg
position: "Assistant Professor of STEM Education"
address: "1328 Beacon Hill Ln., Knoxville, TN 37919"
phone: "+1 865-236-1445"
www: joshuarosenberg.com
email: "jmrosen48@gmail.com"
twitter: jrosenberg6432
github: jrosen48
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::moderncv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tidyverse)
```

# Highlights

- **Researcher** drawing on and contributing to data science education, the learning sciences, and computing education
- **Author** of more than 40 journal articles in leading educational research and social science journals
- **Principal Investigator** for more than 4M in federally-funded research projects, including an National Science Foundation-funded project on combining human and machine-driven approaches to the analysis of classroom audiovisual data
- **Mentor** to graduate and undergraduate students in computer science, data science, and education
- **Data scientist** developing open-source software for learners and researchers

# Professional Affiliation

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Assistant Professor of STEM Education", "2018-Present", "University of Tennessee", "Knoxville, TN",
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Education

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "PhD, Educational Psychology & Educational Technology", "2018", "Michigan State University", "East Lansing, MI",
  "MA, Educational Technology", "2012", "Michigan State University", "East Lansing, MI",
  "BS, Biology", "2010", "University of North Carolina, Asheville", "Asheville, NC",
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Awards and Fellowships (Selected)

```{r}
tribble(
  ~ Award, ~ Year, ~ Institution, ~ Where,
  "Early Career Award", "2022", "Technology as an Agent of Change Special Interest Group", "American Educational Research Association",
  "Best Poster Award", "2021", "Best Poster Award", "International Conference on Educational Data Mining",
  "Fellowship", "2021-2022", "Open Educational Resources (OER) Research", "William and Flora Hewlett Foundation"
) %>% 
  detailed_entries(Award, Year, Institution, Where)
```

# Grants (Selected)

```{r}
tribble(
  ~Role,  ~Years, ~ Title, ~Funder,
  "PI","2019-2022",  "Understanding the development of interest in computer science: An experience sampling approach", "NSF ($348,688)",
  "Co-PI", "2019-2022", "Advancing computational grounded theory for audiovisual data from STEM classrooms", "NSF ($1,313,855)",
) %>% 
    detailed_entries(Role, Years,  Title, Funder)
```

# Publications (Selected)

```{r}
bibliography_entries("publications.bib") %>%
  arrange(issued)
```

