---
title: How many groups of Star Wars characters are there? R-squared and cross-validation
  approaches
author: ~
date: '2017-07-02'
slug: how-many-groups-of-star-wars-characters-are-there-r-squared-and-cross-validation-approaches
tags: [cluster analysis, r, prcr]
description: ''
---

# Background

How many groups, or types, of Star Wars characters are there? I've been wanting to use the `starwars` dataset built-in to the `dplyr` package, and at the same time, have been working hard on an R package to carry out an analysis suited to doing this. Part of the challenge of using the approach in this R package is determining how groups groups there are. 

Many approaches (Latent Profile Analysis, for example) use Maximum Likelihood estimation (while the approach I've developed uses a two-step cluster analysis based around the geometric (and algebraic) idea of "distance", or how close (similar) observations are). This is easy enough when we're talking about something like length. If something is `4` long and another thing `8`, then what is there distance (`4`!)? When we're talking about more than just length - say, length and width - then it's the exact same idea, except the distance represents how far two things are across both measures - length and width. 

But back to groups of Star Wars characters. How many are there? Let's see what data we have:

```{r, message = F}
library(dplyr)

starwars
```

It looks like we only have three measures that are numbers (`height`, `mass`, and `birth_year`) - though there are others we could possibly turn into numbers (maybe), and there are other approaches (Latent Class Analysis) that can deal with non-numeric measures (such as `hair_color`). But we'll have to stick to the three measures that are numbers, for better or worse, for now. 

# R<sup>2</sup>

Let's first take a look at the plot of R<sup>2</sup> values, which are obtained from the second of the two steps of the cluster analysis - the k-means step (I say this because there are other, perhaps better, ways to calculate the R-squared values, such as from a MANOVA). 

We just list the name of the data and the variables we would like to use. Since `birth_year` is on a very different metric than the other two variables, we'll set `to_scale` and `to_center` to `TRUE`. We'll also return a table, instead of a plot. 

```{r}
library(prcr)

plot_r_squared(starwars, height, mass, birth_year, to_scale = TRUE, to_center = TRUE, r_squared_table = T)
```

Ooh! Not good. Before the second of the two steps settled on the groups, it ended up with a group with no observations. This is probably in part the result of a small sample, and possibly attributable to the measures we used - and maybe some missing data for some of the measures. Let's take a look at the data:

```{r}
starwars_ss <- select(starwars, height, mass, birth_year)
skimr::skim(starwars_ss)
```

It looks like the `birth_year` is missing for a lot - 44 - of the observations for the 87 Star Wars characters we have. We're down to the bare-bones number of measures, but let's try with just height and mass. We probably don't need to scale the data.

```{r}
plot_r_squared(starwars, height, mass, to_scale = TRUE, to_center = TRUE, r_squared_table = T)
```

That's better - in a sense. We have two, three, and six groups solutions. I wouldn't trust the six group solution very much. The R<sup>2</sup> value does increase substantialy between two and three groups. This suggests maybe there are three groups (when we use just the measures for weight and mass). 

# Groups

```{r, message = F}
two_profiles <- create_profiles(starwars, height, mass, n_profiles = 2, to_scale = TRUE, to_center = TRUE)
plot(two_profiles)

three_profiles <- create_profiles(starwars, height, mass, n_profiles = 3, to_scale = TRUE, to_center = TRUE)
plot(three_profiles)
```

## The third group: Massive, not so tall

It looks like there is one very massive (literally) observation that makes up one profile in both the two and three profile solutions. Who is it?

```{r}
three_profiles$.data %>% 
    filter(cluster == 3) %>% 
    knitr::kable()
```

Jabba. Of course. It looks like with two or three groups, Jabba ends up in one cluster. 

## The second group: Less massive, small height

What about the seven - who seem to be less massive and with a small height - in the second group?

```{r}
three_profiles$.data %>% 
    filter(cluster == 2) %>% 
    knitr::kable()
```

These seem to be droids, Yoda, and some other tiny characters.

## (Some from) the first group: Above average height, below average mass

The 51 in the first group, with slightly above average height, and slightly below average mass? It's a big group, so here are just the first six, with a lot of familiar characters:

```{r}
three_profiles$.data %>% 
    filter(cluster == 1) %>% 
    head() %>% 
    knitr::kable()
```

# Cross-validation

The other technique for determining the number of groups, cross-validation, may be folly because of how it works: Split the data into two, and see how well groups in one half can be reproduced in the other. This may be a problem due to the Jabba-group.

We'll use the same arguments except for `plot_r_squared`, which we don't need, and for one argument, `n_profiles`, for how many groups we want to cross-validate the groupings for (we have to deal with complete cases, which is what the first two lines are for), for the three group solution:


```{r, eval = F}
starwars_ss <- starwars_ss[complete.cases(starwars_ss), ]
cross_validate(starwars_ss, height, mass, n_profiles = 2, to_scale = TRUE, to_center = TRUE)
```

Not pretty. Convergence issues galore (I decided not to print the messages because there were so many). The Fleiss' Kappa was close to 0; the percentage agreement 0.61.

# Conclusion

Looking at `height` and `weight`, we seem to be able to identify three broad groups of Star Wars characters. However, we shouldn't have a ton of confidence in howe well these groups generalize to all Star Wars characters: Our sample is small, the measures we could use were limited, and our cross-validation did not provide us with much evidence to back up our three distinct groups. 

On the other hand, we did have a starting point for how many groups to look for from our R<sup>2</sup> values, which was good, and the groups seem interpretable on the basis of those characters in our three groups.

# Try it out

The `prcr` package used to create the groups and calculate the R<sup>2</sup> values is available in R using `install.packages("prcr")`. An in-development version with the function for cross-validation is available using the following two commands (if you have devtools installed already then only the second command is needed:

```{r, eval = F}
install.packages("devtools")
devtools::install_github("jrosen48/prcr")
```

Thanks and credit to [Rebecca Steingut](https://motivationlab.wordpress.com/) now at Teachers's College -  Columbia University for contributing to the in-development version of the package and the cross validation strategy implemented in it. 