---
title: Basic {sf} functionality for analyzing and plotting geospatial data
author: Joshua Rosenberg
date: '2020-06-08'
slug: basic-sf-functionality-for-analyzing-and-plotting-geospatial-data
categories: [r]
tags: []
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(message = FALSE)
```

This post has some information on using the 
[{sf} (or simple features) R package](https://r-spatial.github.io/sf/articles/sf1.html)
for analyzing and plotting geospatial data. It is mostly for myself to reference
in the future. A lot of this is from the [*Geocomputation With R* book](https://geocompr.robinlovelace.net/).

## Loading packages, setting up

Load the {sf} and {tidyverse} packages:

```{r}
library(sf)
library(tidyverse)
```

## Creating test data

Create some test data. Here are five longitudes and latitudes. Two are special

- the fourth pair correspond to Knoxville, TN, USA
- the fifth pair corresponds to Lansing, MI, USA

The others are random (and are not within the United States).

```{r}
points <- data.frame(lon = c(-65,-60, -55, -83.92, -84.55), lat = c(50, 45, 40, 35.96, 42.73))
```

The following function creates an **sf** object (telling R which columns within
`points` are the longitude and latitude).

```{r}
points <- st_as_sf(points, coords = c("lon", "lat"))
```

*One thing I'm uncertain about is how {sf} knows that "lon" and "lat" have special meanings;
what if, instead, I had named these something else? Would {sf} know how to prepare to plot them in
the correct way?* Let me know if you know the answer.

## Using sf objects as data frames

Attributes can be added to the **sf** object. The  book suggests
thinking of **sf** objects as containing spatial (e.g. lon, lat) data, and 
attributes; in addition to/because of this, **sf** objects can be treated as 
data frames (in many/most cases). 

Here, I add names to the five points, using indexing as if the object was a data frame.

```{r}
points$name <- c("random_place_c", "random_place_b", "random_place_c", "knoxville", "lansing")
```

It's possible to drop the spatial/geometric part (class) of the object, which can be
useful in some cases; the book mentions that the spatial part is "sticky" and can
be difficult to ignore, even when you use the object as a data frame.

```{r}
st_drop_geometry(points)
```

Here, it is easy to use a **dplyr** verb to remove the random places:

```{r}
points %>% 
    filter(!str_detect(name, "random_place"))
```

## Spatial operations

This step was the big leap for me, but it also illustrates why it can be useful
to treat spatial data in a special way.

U.S. states can be represented with a shape representing each state's boundaries.

There are a number of R packages that provide these shapes in **sf** format. I
am using **USAboundaries** (mostly because I found it first).

```{r}
library(USAboundaries) 

USA <- us_boundaries()
```

I'm not fully grasping the details, but coordinate reference systems (CRSs)
link spatial coordinates (e.g., lon, lat) to the three-dimensional shape of the
Earth. There are short codes for these; a common CRS is 4269 (I have more to 
learn about this). This is important for the next step, the spatial operation.

```{r}
USA <- st_set_crs(USA, "4269")
```

`st_within()` provides information about relationships between pairs of **sf** objects. 
There are many spatial operations in addition to `st_within()`, including `st_intersects()`
and `st_overlaps()`.

```{r}
points_within_USA <- st_within(points, USA)

points_within_USA
```

This can be turned into an integer vector (or another kind of vector, like a 
logical vector).

```{r}
as.integer(points_within_USA)
```

This can be used to index `USA` to see which state the points in `dd` are within:

```{r}
USA[as.integer(points_within_USA), ] %>% 
    filter(!is.na(name))
```

It may be easier to specify the column with the name of the state, too

```{r}
USA[as.integer(points_within_USA), "name"] %>% 
    filter(!is.na(name))
```

## Plotting

**sf** objects can be plotted with ggplot2 using the *sf geom*; here is a plot
of the state boundaries and the five points.

I'll plot only the continental U.S.

```{r}
ggplot() +
    geom_sf(data = filter(USA, name != "Hawaii" & name != "Alaska")) +
    geom_sf(data = points)
```

The projection (and theme) of the map can certainly be improved; this is more to 
demonstrate that {sf} objects can be plotted with {ggplot2}.

Many packages make it easier to plot spatial data of interest. Here is a plot
of all of the school districts in Tennessee.

```{r}
library(leaidr)
library(mapproj)

# only needed the first time
# lea_get(path = "./leaid_sh")

# 47 is the FIPS code for TN
tn <- lea_prep(path = "./leaid_sh", fips = "47")

tn_df <- ggplot2::fortify(tn)

tn_map <-
    ggplot() +
    geom_path(data = tn, 
              aes(x = long, y = lat, group = group),
              color = "gray", size = .2) +
    theme_void()

tn_map_projected <- 
    tn_map +
    coord_map()

print(tn_map_projected)
```

(h/t [Isabella Velásquez](https://ivelasq.rbind.io/)
for making this so seamless (even with fairly ginormous files) to do with the [{leaidr} package](https://github.com/ivelasq/leaidr)).

## *fin*

There is a lot more (for me) to learn; this is just a start.

Resources: 

- [Geocomputation with R](https://geocompr.robinlovelace.net/)
- [Drawing beautiful maps programmatically with R, sf, and ggplot2](https://www.google.com/search?q=ggplot2+sf&rlz=1C5CHFA_enUS724US724&oq=ggplot2+sf&aqs=chrome..69i57j0l5.2310j0j7&sourceid=chrome&ie=UTF-8)
- [r-spatial](https://www.r-spatial.org/)
