%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode
% Awesome CV LaTeX Template for CV/Resume
%
% This template has been downloaded from:
% https://github.com/posquit0/Awesome-CV
%
% Author:
% Claud D. Park <posquit0.bj@gmail.com>
% http://www.posquit0.com
%
%
% Adapted to be an Rmarkdown template by Mitchell O'Hara-Wild
% 23 November 2018
%
% Template license:
% CC BY-SA 4.0 (https://creativecommons.org/licenses/by-sa/4.0/)
%
%-------------------------------------------------------------------------------
% CONFIGURATIONS
%-------------------------------------------------------------------------------
% A4 paper size by default, use 'letterpaper' for US letter
\documentclass[11pt, a4paper]{awesome-cv}

% Configure page margins with geometry
\geometry{left=1.4cm, top=.8cm, right=1.4cm, bottom=1.8cm, footskip=.5cm}

% Specify the location of the included fonts
\fontdir[fonts/]

% Color for highlights
% Awesome Colors: awesome-emerald, awesome-skyblue, awesome-red, awesome-pink, awesome-orange
%                 awesome-nephritis, awesome-concrete, awesome-darknight

\colorlet{awesome}{awesome-red}

% Colors for text
% Uncomment if you would like to specify your own color
% \definecolor{darktext}{HTML}{414141}
% \definecolor{text}{HTML}{333333}
% \definecolor{graytext}{HTML}{5D5D5D}
% \definecolor{lighttext}{HTML}{999999}

% Set false if you don't want to highlight section with awesome color
\setbool{acvSectionColorHighlight}{true}

% If you would like to change the social information separator from a pipe (|) to something else
\renewcommand{\acvHeaderSocialSep}{\quad\textbar\quad}

\def\endfirstpage{\newpage}

%-------------------------------------------------------------------------------
%	PERSONAL INFORMATION
%	Comment any of the lines below if they are not required
%-------------------------------------------------------------------------------
% Available options: circle|rectangle,edge/noedge,left/right

\name{Joshua}{Rosenberg}

\position{Assistant Professor, STEM Education}

\email{\href{mailto:jmrosen48@gmail.com}{\nolinkurl{jmrosen48@gmail.com}}}
\homepage{joshuamrosenberg.com}
\github{jrosen48}
\twitter{jrosenberg6432}

% \gitlab{gitlab-id}
% \stackoverflow{SO-id}{SO-name}
% \skype{skype-id}
% \reddit{reddit-id}


\usepackage{booktabs}

% Templates for detailed entries
% Arguments: what when with where why
\usepackage{etoolbox}
\def\detaileditem#1#2#3#4#5{%
\cventry{#1}{#3}{#4}{#2}{\ifx#5\empty\else{\begin{cvitems}#5\end{cvitems}}\fi}\ifx#5\empty{\vspace{-4.0mm}}\else\fi}
\def\detailedsection#1{\begin{cventries}#1\end{cventries}}

% Templates for brief entries
% Arguments: what when with
\def\briefitem#1#2#3{\cvhonor{}{#1}{#3}{#2}}
\def\briefsection#1{\begin{cvhonors}#1\end{cvhonors}}

\providecommand{\tightlist}{%
	\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%------------------------------------------------------------------------------



\begin{document}

% Print the header with above personal informations
% Give optional argument to change alignment(C: center, L: left, R: right)
\makecvheader

% Print the footer with 3 arguments(<left>, <center>, <right>)
% Leave any of these blank if they are not needed
% 2019-02-14 Chris Umphlett - add flexibility to the document name in footer, rather than have it be static Curriculum Vitae
\makecvfooter
  {September 2020}
    {Joshua Rosenberg~~~·~~~Curriculum Vitae}
  {\thepage}


%-------------------------------------------------------------------------------
%	CV/RESUME CONTENT
%	Each section is imported separately, open each file in turn to modify content
%------------------------------------------------------------------------------



\hypertarget{highlights}{%
\section{Highlights}\label{highlights}}

\begin{itemize}
\tightlist
\item
  Experienced computational social scientist and data scientist
\item
  Develop of open-source software, including three R packages on CRAN
\item
  Manager of an educational data science research group with three NSF grants
\end{itemize}

\hypertarget{experience}{%
\section{Experience}\label{experience}}

\detailedsection{\detaileditem{Assistant Professor, STEM Education}{2018-2020}{University of Tennessee}{Knoxville, TN}{\empty}}

\hypertarget{education}{%
\section{Education}\label{education}}

\detailedsection{\detaileditem{PhD, Educational Psychology and Educational Technology}{2018}{Michigan State University}{East Lansing, MI}{\empty}\detaileditem{MA, Education}{2014}{Michigan State University}{East Lansing, MI}{\empty}\detaileditem{BS, Biology}{2012}{University of North Carolina, Asheville}{Asheville, NC}{\empty}}

\hypertarget{external-funding}{%
\section{External Funding}\label{external-funding}}

\hypertarget{grants}{%
\section{Grants}\label{grants}}

2019-2021, Principle Investigator (PI), \emph{Understanding the development of interest in computer science: An experience sampling approach} (\$346,688). National Science Foundation {[}NSF{]}. \url{http://picsul.utk.edu/} (NSF Grant No.~\href{https://www.nsf.gov/awardsearch/showAward?AWD_ID=1937700\&HistoricalAwards=false}{1937700})

2019-2021, Co-PI, \emph{CS for Appalachia: A research-practice partnership for integrating computer science into East Tennessee schools} (\$252,453; \emph{PI}: Lynn Hodge, University of Tennessee, Knoxville). NSF. (NSF Grant No.~\href{https://www.nsf.gov/awardsearch/showAward?AWD_ID=1923509\&HistoricalAwards=false}{1923509})

2019-2022, Co-PI, \emph{Advancing computational grounded theory for audiovisual data from STEM classrooms} (\$1,313,855; \emph{PI}: Christina Krist, University of Illinois Urbana-Champaign; University of Tennessee, Knoxville subcontract: \$101,469). NSF. \url{https://tca2.education.illinois.edu/} (NSF Grant No.~\href{https://www.nsf.gov/awardsearch/showAward?AWD_ID=1920796\&HistoricalAwards=false}{1920796})

2019-2020, PI, \emph{Planting the seeds for computer science education in East Tennessee through a research-practice partnership} (\$13,200). Community Engaged Research Seed Program, University of Tennessee, Knoxville.

\hypertarget{select-publications}{%
\section{Select Publications}\label{select-publications}}

\hangindent=2em
Estrellado, R. A., Freer, E. A., Mostipak, J., Rosenberg, J. M., \& Velásquez, I. C. (in press). Data science in education using R. London, England: Routledge. Nb. All authors contributed equally. \url{http://www.datascienceineducation.com/}

\hangindent=2em
Rosenberg, J. M., Reid, J., Dyer, E., Koehler, M. J., Fischer, C., \& McKenna, T. J. (in press). Idle chatter or compelling conversation? The potential of the social media-based \#NGSSchat network as a support for science education reform efforts. \emph{Journal of Research in Science Teaching}.

\hangindent=2em
Anderson, D. J., Rowley, B., Stegenga, S., Irvin, P. S., \& Rosenberg, J. M. (advance online publication). Evaluating content-related validity evidence using a text-based, machine learning procedure. Educational Measurement: Issues and Practice. \url{https://onlinelibrary.wiley.com/doi/abs/10.1111/emip.12314}

\hangindent=2em
Greenhalgh, S. P., Rosenberg, J. M., Koehler, M. J., Akcaoglu, M., \& Staudt Willet, K. B. (2020). Identifying multiple learning spaces within a single teacher-focused Twitter hashtag. Computers \& Education, 148(4). \url{https://doi.org/10.1016/j.compedu.2020.103809}

\hangindent=2em
Xu, R., Frank, K. A., Maroulis, S., \& Rosenberg, J. M. (2019). konfound: Command to quantify robustness of causal inferences. The Stata Journal, 19(3), 523-550.

\hangindent=2em
Rosenberg, J. M., van Lissa, C. J., Beymer, P. N., Anderson, D. J., Schell, M. J. \& Schmidt, J. A. (2019). tidyLPA: Easily carry out Latent Profile Analysis (LPA) using open-source or commercial software {[}R package{]}. \url{https://data-edu.github.io/tidyLPA/}

\emph{See all publications here}: \url{https://joshuamrosenberg.com/about/\#publications}

\hypertarget{select-presentations}{%
\section{Select Presentations}\label{select-presentations}}

\hangindent=2em
Rosenberg, J. M., Qinyun, L., Xu, R., Maroulis, S., \& Frank, K. A. (July, 2020). \emph{The konfound R package and Shiny app for robustness analysis}. Presentation at the useR conference, St.~Louis, MO.

\hangindent=2em
Rosenberg, J. M., Beymer, P. N., Phun, V., Schmidt, J. A. (2020, April). Sources of variability for students' engagement in science: Findings from a cross-classified, multivariate modeling approach. In P. N. Beymer, D. K. Benden, \& M. L. Bernacki (Chairs), \emph{Affordances and modeling of intensive data}. Symposium conducted at the American Educational Research Association Annual Meeting, San Francisco, CA.

\hangindent=2em
Rosenberg, J. M, Beymer, P. N., Houslay, T. M., \& Schmidt, J. A. (2019, April). Using a multivariate, multi-level model to understand how youths' in-the-moment engagement predicts changes in youths' interest. In M. Bernacki, A. Kaplan, and L. Linnenbrink-Garcia (Chairs), \emph{Embracing and modeling the complex dynamics of motivation and engagement: Contextual, temporal, dynamic, and systematic}. Symposium conducted at the Annual Meeting of the American Educational Research Association, Toronto, CA.

\emph{See all presentations here}: \url{https://joshuamrosenberg.com/about/\#presentations}

\hypertarget{software-developed}{%
\section{Software Developed}\label{software-developed}}

\hypertarget{r-packages-on-the-comprehensive-r-archive-network-cran}{%
\subsection{R packages on the Comprehensive R Archive Network (CRAN)}\label{r-packages-on-the-comprehensive-r-archive-network-cran}}

Rosenberg, J. M., van Lissa, C. J., Beymer, P. N., Anderson, D. J., Schell, M. J. \& Schmidt, J. A. (2019). \emph{tidyLPA: Easily carry out Latent Profile Analysis (LPA) using open-source or commercial software} {[}R package{]}. \url{https://data-edu.github.io/tidyLPA/}

Rosenberg, J. M., Xu, R., \& Frank, K. A. (2019). \emph{konfound: Quantify the robustness of causal inferences} {[}R package{]}. \url{https://jrosen48.github.io/konfound/}

Rosenberg, J. M., Schmidt, J. A., Beymer, P. N., \& Steingut, R. (2018). \emph{prcr: Person-Centered Analysis} {[}R package{]}. \url{https://CRAN.R-project.org/package=prcr}

\hypertarget{r-packages-on-github}{%
\subsection{R packages on GitHub}\label{r-packages-on-github}}

Estrellado, R. A., Bovee, E. A., Mostipak, J., Rosenberg, J. M., \& Velásquez, I. C. (2019). \emph{dataedu: Package for Data Science in Education Using R}. \url{https://github.com/data-edu/dataedu}

Anderson, D. Heiss, A., and Rosenberg, J. M. (2019). \emph{equatiomatic: Transform Models into LaTeX Equations.} \url{https://github.com/datalorax/equatiomatic}

\hypertarget{interactive-web-application}{%
\subsection{Interactive Web Application}\label{interactive-web-application}}

Rosenberg, J. M., Xu, R., \& Frank, K. A. (2019). \emph{Konfound-It!: Quantify the robustness of causal inferences.} \url{http://konfound-it.com}.

\hypertarget{web-application}{%
\subsection{Web Application}\label{web-application}}

Lishinski, A., \& Rosenberg, J. M. (2019). \emph{Short message survey: An open-source, text-message based application for the experience sampling method.} \url{https://github.com/picsul/short-message-survey}

\end{document}
