---
name: Joshua
surname: Rosenberg
position: "Assistant Professor, STEM Education"
www: joshuamrosenberg.com
email: "jmrosen48@gmail.com"
twitter: jrosenberg6432
github: jrosen48
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::awesomecv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

# Highlights

 * Experienced computational social scientist and data scientist
 * Develop of open-source software, including three R packages on CRAN
 * Manager of an educational data science research group with three NSF grants

# Experience

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Assistant Professor, STEM Education", "2018-2020", "University of Tennessee", "Knoxville, TN",
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "PhD, Educational Psychology and Educational Technology", "2018", "Michigan State University", "East Lansing, MI",
  "MA, Education", "2014", "Michigan State University", "East Lansing, MI",
  "BS, Biology", "2012", "University of North Carolina, Asheville", "Asheville, NC",
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# External Funding

# Grants

2019-2021, Principle Investigator (PI), *Understanding the development of interest in computer science: An experience sampling approach* ($346,688). National Science Foundation [NSF]. http://picsul.utk.edu/ (NSF Grant No. [1937700](https://www.nsf.gov/awardsearch/showAward?AWD_ID=1937700&HistoricalAwards=false))

2019-2021, Co-PI, *CS for Appalachia: A research-practice partnership for integrating computer science into East Tennessee schools* ($252,453; *PI*: Lynn Hodge, University of Tennessee, Knoxville). NSF. (NSF Grant No. [1923509](https://www.nsf.gov/awardsearch/showAward?AWD_ID=1923509&HistoricalAwards=false))

2019-2022, Co-PI, *Advancing computational grounded theory for audiovisual data from STEM classrooms* ($1,313,855; *PI*: Christina Krist, University of Illinois Urbana-Champaign; University of Tennessee, Knoxville subcontract: $101,469). NSF. https://tca2.education.illinois.edu/ (NSF Grant No. [1920796](https://www.nsf.gov/awardsearch/showAward?AWD_ID=1920796&HistoricalAwards=false))

2019-2020, PI, *Planting the seeds for computer science education in East Tennessee through a research-practice partnership*  ($13,200). Community Engaged Research Seed Program, University of Tennessee, Knoxville.

# Select Publications

\hangindent=2em
Estrellado, R. A., Freer, E. A., Mostipak, J., Rosenberg, J. M., & Velásquez, I. C. (in press). Data science in education using R. London, England: Routledge. Nb. All authors contributed equally. http://www.datascienceineducation.com/

\hangindent=2em
Rosenberg, J. M., Reid, J., Dyer, E., Koehler, M. J., Fischer, C., & McKenna, T. J. (in press). Idle chatter or compelling conversation? The potential of the social media-based #NGSSchat network as a support for science education reform efforts. *Journal of Research in Science Teaching*.

\hangindent=2em
Anderson, D. J., Rowley, B., Stegenga, S., Irvin, P. S., & Rosenberg, J. M. (advance online publication). Evaluating content-related validity evidence using a text-based, machine learning procedure. Educational Measurement: Issues and Practice. https://onlinelibrary.wiley.com/doi/abs/10.1111/emip.12314

\hangindent=2em
Greenhalgh, S. P., Rosenberg, J. M., Koehler, M. J., Akcaoglu, M., & Staudt Willet, K. B. (2020). Identifying multiple learning spaces within a single teacher-focused Twitter hashtag. Computers & Education, 148(4). https://doi.org/10.1016/j.compedu.2020.103809

\hangindent=2em
Xu, R., Frank, K. A., Maroulis, S., & Rosenberg, J. M. (2019). konfound: Command to quantify robustness of causal inferences. The Stata Journal, 19(3), 523-550.

\hangindent=2em
Rosenberg, J. M., van Lissa, C. J., Beymer, P. N., Anderson, D. J., Schell, M. J. & Schmidt, J. A. (2019). tidyLPA: Easily carry out Latent Profile Analysis (LPA) using open-source or commercial software [R package]. https://data-edu.github.io/tidyLPA/

*See all publications here*: https://joshuamrosenberg.com/about/#publications

# Select Presentations

\hangindent=2em
Rosenberg, J. M., Qinyun, L., Xu, R., Maroulis, S., & Frank, K. A. (July, 2020). *The konfound R package and Shiny app for robustness analysis*. Presentation at the useR conference, St. Louis, MO.

\hangindent=2em
Rosenberg, J. M., Beymer, P. N., Phun, V., Schmidt, J. A. (2020, April). Sources of variability for students' engagement in science: Findings from a cross-classified, multivariate modeling approach. In P. N. Beymer, D. K. Benden, & M. L. Bernacki (Chairs), *Affordances and modeling of intensive data*. Symposium conducted at the American Educational Research Association Annual Meeting, San Francisco, CA.

\hangindent=2em
Rosenberg, J. M, Beymer, P. N., Houslay, T. M., & Schmidt, J. A. (2019, April). Using a multivariate, multi-level model to understand how youths’ in-the-moment engagement predicts changes in youths’ interest. In M. Bernacki, A. Kaplan, and L. Linnenbrink-Garcia (Chairs), *Embracing and modeling the complex dynamics of motivation and engagement: Contextual, temporal, dynamic, and systematic*. Symposium conducted at the Annual Meeting of the American Educational Research Association, Toronto, CA. 

*See all presentations here*: https://joshuamrosenberg.com/about/#presentations

# Software Developed

## R packages on the Comprehensive R Archive Network (CRAN)

Rosenberg, J. M., van Lissa, C. J., Beymer, P. N., Anderson, D. J., Schell, M. J. & Schmidt, J. A. (2019). *tidyLPA: Easily carry out Latent Profile Analysis (LPA) using open-source or commercial software* [R package]. https://data-edu.github.io/tidyLPA/

Rosenberg, J. M., Xu, R., & Frank, K. A. (2019). *konfound: Quantify the robustness of causal inferences* [R package]. https://jrosen48.github.io/konfound/

Rosenberg, J. M., Schmidt, J. A., Beymer, P. N., & Steingut, R. (2018). *prcr: Person-Centered Analysis* [R package]. https://CRAN.R-project.org/package=prcr

## R packages on GitHub

Estrellado, R. A., Bovee, E. A., Mostipak, J., Rosenberg, J. M., & Velásquez, I. C. (2019). *dataedu: Package for Data Science in Education Using R*. https://github.com/data-edu/dataedu

Anderson, D. Heiss, A., and Rosenberg, J. M. (2019). *equatiomatic: Transform Models into LaTeX Equations.* https://github.com/datalorax/equatiomatic

## Interactive Web Application

Rosenberg, J. M., Xu, R., & Frank, K. A. (2019). *Konfound-It!: Quantify the robustness of causal inferences.* http://konfound-it.com.

## Web Application

Lishinski, A., & Rosenberg, J. M. (2019). *Short message survey: An open-source, text-message based application for the experience sampling method.* https://github.com/picsul/short-message-survey