---
title: "Comparing MPLUS and MCLUST output"
author: "Joshua Rosenberg"
date: "2017-09-14"
output: html_document
categories: [r]
---

### Note

An in-development `R` package to carry out LPA and this analysis is available here: [https://github.com/jrosen48/tidymixmod](https://github.com/jrosen48/tidymixmod).

# Introduction

At present, MPlus is a widely-used tool to carry out Latent Profile Analysis, and there does not seem to be a widely-accepted or used way to carry out Latent Profile Analysis in R. This compares output from MPlus to output from the R package MCLUST, which is accessed through the package tidymixmod which I started to work on to make estimating common models used for LPA easier to carry out using MCLUST. 

### Links to resources:

- [MPlus](https://www.statmodel.com/)
- [MPlusAutomation](https://www.statmodel.com/usingmplusviar.shtml) (this package uses MPlus; MPlus has to be purchased and installed for it to work)
- [MCLUST](http://www.stat.washington.edu/mclust/)
- [tidymixmod](https://github.com/jrosen48/tidymixmod) (this is the package we're currently working on; it provides an interface to MCLUST)

# Loading, setting up

First, let's load a few packages and setup.

```{r, message = F}
library(tidyverse)
library(MplusAutomation)
library(stringr)
# if tidymixmod is not installed, can use the line below to install
# devtools::install_github("jrosen48/tidymixmod")
library(tidymixmod)

the_dir <- "/Users/joshuarosenberg/Dropbox/5_Professional/homepage/source/static/_media/data/"

```

Next, we'll write a few functions to view MPlus summary statistics from the output.

```{r}
extract_mplus_summary <- function(x) {
    x$summaries[c("LL", "BIC", "Entropy")]
}

extract_mplus_paramaters <- function(x) {
    y <- x$parameters[[1]]
    y <- y[-nrow(y), ]
    dplyr::select(y, param_name = paramHeader, var_name = param, est, se, class = LatentClass)
}

extract_means <- function(x) {
    calculate_centroids_mclust(x) %>%
        rownames_to_column("var_name") %>%
        rename(class_1 = Profile1, class_2 = Profile2) %>%
        mutate(param_name = "Means",
               class1 = round(class_1, 3),
               class2 = round(class_2, 3)) %>%
        select(param_name, var_name, class_1, class_2)
}
```

Here we are running MPlus models using MPlusAutomation. The specification of each of the three models is available from the post [here](https://jrosen48.github.io/blog/using-mplus-from-r/). 

```{r, results = "hide"}
runModels(paste0(the_dir, "2-iris-LPA_means.inp"))
runModels(paste0(the_dir, "2-iris-LPA_means_correlated.inp"))
runModels(paste0(the_dir, "2-iris-LPA_means_correlated_free_variances.inp"))

m1_mplus_out <- readModels(paste0(the_dir, "2-iris-LPA_means.out"))
m2_mplus_out <- readModels(paste0(the_dir, "2-iris-LPA_means_correlated.out"))
m3_mplus_out <- readModels(paste0(the_dir, "2-iris-LPA_means_correlated_free_variances.out"))

m1_mplus <- m1_mplus_out %>%
    extract_mplus_paramaters() %>%
    mutate(class = paste0("class_", class)) %>%
    select(-se) %>%
    spread(class, est)

m2_mplus <- m2_mplus_out %>%
    extract_mplus_paramaters() %>%
    mutate(class = paste0("class_", class)) %>%
    select(-se) %>%
    spread(class, est)

m3_mplus <- m3_mplus_out %>%
    extract_mplus_paramaters() %>%
    mutate(class = paste0("class_", class)) %>%
    select(-se) %>%
    spread(class, est)
```

Here we are running MCLUST models using the R package [tidymixmod](https://github.com/jrosen48/tidymixmod), which interfaces to the R package MCLUST.

```{r}
m1_mclust <- create_profiles_mclust(iris,
                                    Sepal.Length, Sepal.Width, Petal.Length, Petal.Width,
                                    n_profiles = 2, model = 1, to_return = "mclust")
m2_mclust <- create_profiles_mclust(iris,
                                    Sepal.Length, Sepal.Width, Petal.Length, Petal.Width,
                                    n_profiles = 2, model = 2, to_return = "mclust")
m3_mclust <- create_profiles_mclust(iris,
                                    Sepal.Length, Sepal.Width, Petal.Length, Petal.Width,
                                    n_profiles = 2, model = 3, to_return = "mclust")
```

# Comparing MPlus and MCLUST summary statistics

*LL is log-likelihood*

```{r}
list(m1_mplus_out, m2_mplus_out, m3_mplus_out) %>%
    map_df(extract_mplus_summary) %>% 
    mutate(model = paste0("m", 1:3),
           method = "MPlus") %>%
    select(method, model, everything())

list(m1_mclust, m2_mclust, m3_mclust) %>%
    map_df(extract_mclust_summary) %>%
    mutate(model = paste0("m", 1:3),
           method = "MCLUST") %>%
    select(method, model, everything())

```

There is a slight difference in the entropy statistic for the first model (means with constrained variances), possibly due to a rounding error from the MPlus output. 

# Comparing MPlus and MCLUST parameter estimates

#### For m1 (varying means, constrained variances)

```{r}
m1_m <- extract_means(m1_mclust) %>%
    mutate(class_1 = round(class_1, 3),
           class_2 = round(class_2, 3))

m1_c1_v <- extract_variance(m1_mclust, 1)
m1_c2_v <- extract_variance(m1_mclust, 2)

# MPlus output

m1_mplus %>% 
    mutate(model = "MPlus") %>% 
    select(model, everything()) %>% 
    arrange(param_name, desc(class_1))

# MCLUST output

bind_rows(m1_c1_v, m1_c2_v) %>% 
    spread(class, est) %>% 
    bind_rows(m1_m) %>% 
    mutate(model = "MCLUST") %>% 
    select(model, everything()) %>% 
    arrange(param_name, desc(class_1)) %>% 
    mutate(var_name = toupper(var_name),
           var_name = str_sub(var_name, end = 10),
           var_name = str_replace(var_name, "\\.", "_"))
```

These seem to be identical.

#### For m2 (varying means, constrained variances and covariances)

```{r}
m2_m <- extract_means(m2_mclust) %>%
    mutate(class_1 = round(class_1, 3),
           class_2 = round(class_2, 3))

m2_c1_v <- extract_variance(m2_mclust, 1)
m2_c2_v <- extract_variance(m2_mclust, 2)

m2_c1_cv <- extract_covariance(m2_mclust, 1) %>% 
    semi_join(m2_mplus, by = c("param_name", "var_name")) %>% 
    mutate(class = "class_1")

m2_c2_cv <- extract_covariance(m2_mclust, 2) %>% 
    semi_join(m2_mplus, by = c("param_name", "var_name")) %>% 
    mutate(class = "class_2")

m2_cv <- bind_rows(m2_c1_cv, m2_c2_cv) %>% 
    mutate(est = round(est, 3),
           model = "MCLUST") %>% 
    spread(class, est) 

# MPlus output

m2_mplus %>% 
    mutate(model = "MPlus") %>% 
    select(model, everything()) %>% 
    arrange(param_name, desc(class_1)) 

# MCLUST output

bind_rows(m2_c1_v, m2_c2_v) %>% 
    spread(class, est) %>% 
    bind_rows(m1_m) %>% 
    mutate(model = "MCLUST") %>% 
    select(model, everything()) %>% 
    bind_rows(m2_cv) %>% 
    arrange(param_name) %>% 
    arrange(param_name, desc(class_1)) %>% 
    mutate(var_name = toupper(var_name),
           var_name = str_sub(var_name, end = 10),
           var_name = str_replace(var_name, "\\.", "_"))
```

There seem to be a few differences in the hundreths place.

#### For m3 (varying means, freed covariances and covariances)

```{r}
m3_m <- extract_means(m3_mclust) %>%
    mutate(class_1 = round(class_1, 3),
           class_2 = round(class_2, 3))

m3_c1_v <- extract_variance(m3_mclust, 1)
m3_c2_v <- extract_variance(m3_mclust, 2)

m3_c1_cv <- extract_covariance(m3_mclust, 1) %>% 
    semi_join(m3_mplus, by = c("param_name", "var_name")) %>% 
    mutate(class = "class_1")

m3_c2_cv <- extract_covariance(m3_mclust, 2) %>% 
    semi_join(m3_mplus, by = c("param_name", "var_name")) %>% 
    mutate(class = "class_2")

m3_cv <- bind_rows(m3_c1_cv, m3_c2_cv) %>% 
    mutate(est = round(est, 3),
           model = "MCLUST") %>% 
    spread(class, est) 

# MPlus output

m3_mplus %>% 
    mutate(model = "MPlus") %>% 
    select(model, everything()) %>% 
    arrange(param_name, desc(class_1)) 

# MCLUST output

bind_rows(m1_c1_v, m1_c2_v) %>% 
    spread(class, est) %>% 
    bind_rows(m1_m) %>% 
    mutate(model = "MCLUST") %>% 
    select(model, everything()) %>% 
    bind_rows(m3_cv) %>% 
    arrange(param_name) %>% 
    arrange(param_name, desc(class_1)) %>% 
    mutate(var_name = toupper(var_name),
           var_name = str_sub(var_name, end = 10),
           var_name = str_replace(var_name, "\\.", "_"))
```

These seem to be a bit different at the tenths or hundreths decimal point.

# Conclusion

At least for these simple models, the output appears to be identical, with slight differences in the entropy statistic and some of the parameter estimates for the third models.

More complex models **may** be different: 

- MPlus uses random starts to initialize ML, while MClust initializes ML with starts from hierarchical clustering.
- MPlus seems to use restricted MLR to obtain robust standard errors, whereas MClust uses weighted likelihood bootstrap methods; I'm not sure if these are different for more complex models, and so we should check the standard errors.

Future directions include:

- Deciding on a uniform interface for prcr and tidymixmod.
- Merging tidymixmod with [prcr](https://github.com/jrosen48/prcr) (our R package for person-oriented analysis).
