---
title: 'The identical output of lm() and aov() and lm() and t.test() in R'
author: ''
date: '2022-11-21'
slug: ''
categories: []
tags: []
---


This one hardly counts as a blog post; it is more a reference for myself.

For some time, I've known on a conceptual level that a linear regression model (one "fit" using `lm()` in the statistical software and programming language R), ANOVA (`aov()`), and *t*-test (`t.test()`) can be conceptualized as being closely-related---though they're often taught as separate techniques. 

I'm going to add the code now, hoping to return to and expand on this later.

Let's use a data set about penguins.

```{r, message = FALSE}
library(tidyverse)
library(palmerpenguins)

penguins %>% 
    glimpse()
```

## `lm()` and `aov()`

Note how the `*F*-value, *df*, and *p*-value are identical.

```{r}
summary(lm(bill_length_mm ~ island, data = penguins))
```

```{r}
summary(aov(bill_length_mm ~ island, data = penguins))
```

## `lm()` and `t.test()`

Note the *t*-value and its *df* and *p*-value are identical (for the sex variable).

```{r}
summary(lm(bill_length_mm ~ sex, data = penguins))
```

```{r}
t.test(bill_length_mm ~ sex, var.equal = TRUE,data = penguins)
```

I may come back to this later to flesh this out more...
